Lifting CDCL to Relational Domains

### 2016-10-15 domains.tex
learning and domain update

%-------------------------------------------------------------------------------
\subsection{Properties of Abstract Domains}
%-------------------------------------------------------------------------------
%
An important property of a clause learning SAT solver is that each
element of the boolean domain is decomposable into
complementable elements.  This property of the domain is required to
learn elements that help to guide the model search away from the
conflicting region of the search space.
%
Most numerical abstract domains, such as intervals and octagons lack
complements, but they can be represented as intersections of
complementable half-spaces, each of which has a complement. 
%
% We identify specific properties of domain elements necessary for 
% abstract model search and learning in abstractions.  
We formalise this in the sequel.
%
\begin{definition} 
A \emph{meet irreducible} $m$ in a complete lattice 
structure $A$ is an element with the following property.
\[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
%\rmcmt{mention minimum complementable element}
\end{definition}
%
The meet irreducibles in the boolean domain $\mathcal{B}$ for a variable $x$ are 
$x$ and $\neg x$. The meet 
irreducibles in the template polyhedra domain are all elements 
that concretise to half-spaces, i.e.\ they can be represented 
by a single inequality. For the interval domain, these are 
\ $x \leq d$ or $x \geq d$ for constants $d$.

@@

\begin{definition} 
An element $a\in A$ is called \emph{precisely complementable}
iff $\exists \bar{a} \in A: \neg\gamma(\bar{a})=\gamma(a)$, 
i.e.\ there is an element whose negated concretisation equals
the concretisation of $a$.
%m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.
%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
%is the set complement of $\bar{M}$.  
%If every meet irreducible in $A$ is complementable, then $A$ is said to
%be have complementable meet irreducibles.
\end{definition}

>>>>

%-------------------------------------------------------------------------------
\subsection{Properties of Abstract Domains}
%-------------------------------------------------------------------------------
%
An important property of a clause learning SAT solver is that each
non-singleton element of the partial assignment domain is decomposable into
complementable elements.  This property of the domain is required to
learn elements that help to guide the model search away from the
conflicting region of the search space.
\rmcmt{Convention to be followed: In SAT world, partial assignment (PA) is a data structure. Previous ACDL paper 
showed that PA is an abstract domain and called it partial assignment domain. Likewise, in this paper, 
whenever we refer to SAT, we also call partial assignment domain. But When we refer 
to our ACDCL product domain, we call boolean domain and polyhedra domain. We follow this 
convention throughout. This would make a SAT reader happy as well as a Abstract Interpreter not sad !!}
%
Most numerical abstract domains, such as intervals and octagons lack
complements, but they can be represented as intersections of
complementable half-spaces, each of which has a complement. 
%
% We identify specific properties of domain elements necessary for 
% abstract model search and learning in abstractions.  
We formalise this in the sequel.
%
\begin{definition} 
A \emph{meet irreducible} $m$ in a complete lattice 
structure $A$ is an element with the following property.
\[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
%\rmcmt{mention minimum complementable element}
\end{definition}
%
The meet irreducibles in the boolean domain $\mathcal{B}$ for a variable $x$ are 
$x$ and $\neg x$. The meet 
irreducibles in the template polyhedra domain are all elements 
that concretise to half-spaces, i.e.\ they can be represented 
by a single inequality. For the interval domain, these are 
\ $x \leq d$ or $x \geq d$ for constants $d$.

@@

\begin{definition} 
An element $a\in A$ is called \emph{precisely complementable}
iff $\exists \bar{a} \in A: \neg\gamma(\bar{a})=\gamma(a)$, 
i.e.\ there is an element whose negated concretisation equals
the concretisation of $a$.
\rmcmt{$\gamma$ is not defined before and used here for the first time}
%m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.
%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
%is the set complement of $\bar{M}$.  
%If every meet irreducible in $A$ is complementable, then $A$ is said to
%be have complementable meet irreducibles.
\end{definition}

### 2016-10-15 learning.tex
learning and domain update

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$.  
%
\begin{figure}
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\label{conflict}
\end{figure} 
%    
\paragraph{\textbf{Clause Learning in Abstract Lattice}}
A learnt clause $\mathcal{L}$ in propositional solver is obtained 
by negating the reason for conflict.  An asserting clause guarantees 
new dductions using unit rule~\cite{cdcl}.  However, we model a 
learnt clause as abstract deduction transformer, called {\em abstract unit 
transformer (AUnit)}.  For an abstract lattice $A$ and a set of 
countermodels $C \subseteq A$ of $\varphi$ such that \meet{C} does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is $\top$ if $\exists a \in A$ 
such that $a \subseteq \meet{C}$.  However, if $t \in C$ and 
$\forall t' \in C \ {t}$ such that $a \sqsubseteq t'$, $AUnit$ 
infers a valid deduction $\bar{t}$, which directs the search 
away from the conflict region. Else, $AUnit$ returns $\top$.

>>>>

For our example, there exists multiple incomaparable reasons for conflict,
marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
Choosing cut0 yields a learnt clause 
$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$.  
%
\begin{figure}
%\caption{An example of Octagon}\label{octagon}
\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
\label{conflict}
\end{figure} 
%    
\paragraph{\textbf{Clause Learning in Abstract Lattice}}
A learnt clause $\mathcal{L}$ in propositional solver is obtained 
by negating the reason for conflict.  An asserting clause guarantees 
new deductions using unit rule~\cite{cdcl}.  However, we model a 
learnt clause as meet irreducible deduction transformer, 
called {\em abstract unit transformer (AUnit)}.  For an 
abstract lattice $A$ with complementable meet irreducibles 
and a set of countermodels $C \subseteq A$ of $\varphi$ such that \meet{C} does not 
satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is $\top$ if $\exists a \in A$ 
such that $a \subseteq \meet{C}$.  However, if $t \in C$ and 
$\forall t' \in C \\ \{t\}$ such that $a \sqsubseteq t'$, $AUnit$ 
infers a valid deduction $\bar{t}$, which directs the search 
away from the conflict region. Else, $AUnit$ returns $\top$.
